% !TEX root = manual.tex
\chapter{Configuration}

\section{Requirements}

LibreS3 by default listens on ports 8008 and 8443, which need to be available on a given
IP address.

LibreS3 connects to the \SX cluster via \verb|HTTP(S)|. You can run LibreS3 and \SX
on the same or different hosts.

LibreS3 doesn't store state on the local filesystem, when it does need to share
data with other LibreS3 instances (e.g.\ multipart uploads) it will use an \SX
volume for that purpose.
You can run any number of LibreS3 instances, to provide load-balancing or
failover.

\subsection{DNS zone entry}
\label{sec:dnszone}

S3 buckets require a wildcard \verb|A| record pointing to the IP address (1.2.3.4 below) of the
host running LibreS3, for example:

\begin{lstlisting}
*.libres3.example.com. A 1.2.3.4
libres3.example.com. A 1.2.3.4
\end{lstlisting}

If you don't have control over the DNS you'll have to modify the \path{/etc/hosts} file
of each client machine and add a line for each bucket you want to access:

\begin{lstlisting}
libres3.example.com 1.2.3.4
bucket1.libres3.example.com 1.2.3.4
bucket2.libres3.example.com 1.2.3.4
\end{lstlisting}

You can also add \verb|CNAME| for virtual hosting buckets:
\begin{lstlisting}
bucket.example.net CNAME        bucket.example.net.libres3.example.com.
\end{lstlisting}

You can of course just add directly \verb|A| or \verb|AAAA| entries that point
directly to the LibreS3 server, it only matters that
clients send the desired bucket name in the \verb|Host:| header.

\section{Setting up a LibreS3 node}

Setting up LibreS3 is as simple as running the interactive tool \verb+libres3_setup+.
If you provide the path to an existing \SX cluster configuration file created
by \verb+sxsetup+, most of the settings will be done automatically.

Please make sure the default volume replica count setting is less or equal
to the number of nodes in the \SX cluster.

In the examples below we assume that you
have an \SX cluster already up and running and you want to deploy
LibreS3 on \verb|libres3.example.com|.

Example setup with sxsetup.conf:

\begin{lstlisting}
# libres3_setup  --sxsetup-conf /etc/sxserver/sxsetup.conf
Successfully loaded SX configuration from '/etc/sxserver/sxsetup.conf'

S3 (DNS) name: libres3.example.com

Generating default SSL certificate and key in /etc/ssl/certs/libres3.pem and /etc/ssl/private/libres3.key
Generating a 2048 bit RSA private key
.........................................+++
.............+++
writing new private key to '/etc/ssl/private/libres3.key'
-----

S3 HTTPS port: 8443

S3 HTTP port: 8008

Default volume size [use K, M, G and T suffixes]: 100G

Default volume replica count: 1

Generating '/etc/libres3/libres3.conf'
Updating '/etc/libres3/libres3.sample.s3cfg'
Generating '/etc/libres3/libres3.sample.boto'
Updating '/etc/libres3/libres3-insecure.sample.s3cfg'
Generating '/etc/libres3/libres3-insecure.sample.boto'

Do you want to start LibreS3 now? [Y/n] Y
[....] Restarting LibreS3: libres3No libres3_ocsigen found running; none killed.

Loading configuration from /etc/libres3/libres3.conf
Waiting for server to start (5s) ... OK
\end{lstlisting}


Example without sxsetup.conf:
\footnote{you can use \verb|sxadm node --info /var/lib/sxserver/storage/| on the \SX node to find out the required information}

\begin{lstlisting}
# libres3_setup
Admin key or path to key-file: 0DPiKuNIrrVmD8IUCuw1hQxNqZfJ0hlBUgyckAolodd4C/4r4ecY3QAA

SX server IP/DNS name: sx.example.com

SX server HTTPS port: 443

Run as user: nobody

Run as group: nogroup

S3 (DNS) name: libres3.example.com

Generating default SSL certificate and key in /etc/ssl/certs/libres3.pem and /etc/ssl/private/libres3.key
Generating a 2048 bit RSA private key
............................................................+++
............+++
writing new private key to '/etc/ssl/private/libres3.key'
-----

S3 HTTPS port: 8443

S3 HTTP port: 8008

Default volume size [use K, M, G and T suffixes]: 100G

Default volume replica count: 1

Generating '/etc/libres3/libres3.conf'
File '/etc/libres3/libres3.conf' already exists, overwriting
Updating '/etc/libres3/libres3.sample.s3cfg'
Generating '/etc/libres3/libres3.sample.boto'
Updating '/etc/libres3/libres3-insecure.sample.s3cfg'
Generating '/etc/libres3/libres3-insecure.sample.boto'

Do you want to start LibreS3 now? [Y/n] Y
[....] Restarting LibreS3: libres3No libres3_ocsigen found running; none killed.

Loading configuration from /etc/libres3/libres3.conf
Waiting for server to start (5s) ... OK
\end{lstlisting}


To start/stop LibreS3\footnote{LibreS3 and \SX will communicate using TLS by
default. For debugging purposes you can configure \SX with \verb|sxsetup --no-ssl|}:

\begin{lstlisting}
# libres3 start
Starting LibreS3
LibreS3 started successfully
# libres3 status
--- LibreS3 STATUS ---
LibreS3 is running (PID 28245)

--- LibreS3 INFO ---
SSL private key: /etc/ssl/private/libres3.key
LibreS3 logs: /var/log/libres3/
# libres3 stop
Loading configuration from /etc/libres3/libres3.conf
Sending TERM to PID 28245 ... 
Waiting for PID 28245 ...
\end{lstlisting}


If the server doesn't start, please check the log files for details.

That's it --- your LibreS3 cloud storage is already up and running!
You can now connect to it with your favorite S3 client.

\section{Advanced settings}

\subsection{Internal IP addresses}

If you've configured \SX with public and internal IP addresses LibreS3 can only
access \SX via its public IP address, so make sure it is reachable from LibreS3
and that you set the public IP address of \SX in \verb|sx_host|.

\subsection{libres3.conf}

When you run \verb|libres3_setup| it will generate a file
\path{/etc/libres3/libres3.conf}.
You can customize this file and restart LibreS3 for the changes to take effect.

LibreS3 tries to be compatible with the S3 protocol by default, however
sometimes there are features outside of the S3 protocol that would be useful.
These can be configured in \verb|libres3.conf| and all have the \verb|allow_| prefix:
\begin{description}
\item[allow\_list\_all\_volumes (default=false)] only volumes owned by you are
  listed by default, \seeref{sec:listing-buckets}.
\item[allow\_public\_bucket\_index (default=false)] by default you can't list a
  bucket or subdirectory from a browser, \seeref{sec:directory-indexing}.
\item [allow\_volume\_create\_any\_user (default=true)] by default any authenticated
  user can create a bucket that doesn't exist yet. In \SX only admin users can
  perform such an action, and you can configure LibreS3 to provide that
  behaviour, however some S3 applications may not behave correctly anymore.
\end{description}

There are more settings that tune various aspects of LibreS3, an up-to-date
list\footnote{this list doesn't include settings used for debugging/tuning
  low-level aspects} can always be found in the generated \verb|libres3.conf|: each configuration option has
a comment briefly describing its purpose, and settings that are disabled by
default are commented out.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manual"
%%% End:
